---
title: "Project"
output:
  pdf_document: default
  word_document: default
---


```{r}
#Load data  
data<- read.csv("spotify10.csv")

```

```{r}
#Load required packages

list.of.packages <- c("ggplot2","caret","tinytex","rmarkdown","knitr","e1071","randomForest","sqldf","factoextra","nnet","tidyverse","cluster","fpc","shiny","shinyWidgets","neuralnet")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
invisible(lapply(list.of.packages, require, character.only = TRUE))
```

```{r}
#Understand the columns and data types 
head(data)
str(data)

```

```{r}
#Remove Char variables to find correlation matrix
cordata<-data[-1]
cordata<-cordata[-1]
cordata<-cordata[-1]
head(cordata)
```

```{r}
#Unique artists 
length(unique(data$artist))
unique(data$key)

unique(data$time_signature)



```


```{r}

#Find correlation between variables 
cor(cordata)

```
Danceabilty, loudness, acousticness, instrumentalness, valence, sections,time_signature explain the target variable if it's a mainstream hit or a flop song.

key, mode, tempo do not really help the prediction. Once we run a model  to test our hypothesis, we can be sure if there are significant enough to predict the target variable 

```{r}
#check for null values
colSums(is.na(data))

```
No null values 

```{r}
# To check if the data is balanced 
table(data$target)

```
Data is balanced 

```{r}
#further Analysis 

popularartist<-sqldf("Select artist, Count(track) from data where Target=1 group by artist order by Count(Track) desc limit 10")

popularartist

 ggplot(data=popularartist,aes(x=reorder(artist,-`Count(track)`),y=`Count(track)`)) +
  geom_bar(stat="identity", fill="pink")+  geom_text(aes(label=`Count(track)`), vjust=1.6,color="black", size=3.5)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(
   title = 'Artist with maximum number of hit songs from 2010-2019',
    y = 'count of hit songs', x="Artists") 
```
 


```{r}

sqldf("Select artist, Count(track), avg(speechiness), avg(danceability), avg(valence), avg(energy) from data where Target=1 group by artist order by Count(Track) desc limit 10")

```

There is a clear difference is speechiness in hip hop artists such as drake compared to  pop artists such as taylor swift. 

Danceability is relatively high for all the songs on average(above 50), valence(above 35). 
Energy is also really high on average for all these artist but the correlation value was low for energy. Need to run model to check it's significance.

So we say that Danceability and valence are more significant in predicting target variable than speechiness as there values are in difference ranges so they probabilty don't impact the prediction of a mainstream hit song



```{r} 
#Split data into train and test

data1<-cordata
data1$target<-as.factor(data1$target)
d <- createDataPartition(data1$target, p=0.75, list=FALSE)
train <- data1[ d,]
test <- data1[-d,]

nrow(train)
nrow(test)
head(test)
```


``` {r}
# To Find out significant variables in predicting target variable  
#GLM

model1<- glm(target~.,data=train,family='binomial')
summary(model1)

predtest <- predict(model1,newdata =  test,type="response")
t<-table(predtest > 0.5, test$target) 
sum(diag(t)) / sum(t)

```

```{r}

sqldf("select avg(danceability),avg(loudness),avg(energy),avg(speechiness),avg(acousticness),avg(instrumentalness),avg(time_signature) from data where Target=1")

```
All the variables that are highly significant in determining the mainstream hit songs and their average values. 


```{r}
#Models to predict the target variable 

#SVM
train_control <- trainControl(method="cv", number=10) #k-fold cross validation
model2<-svm(target~danceability+energy+loudness+acousticness+instrumentalness+valence+tempo+time_signature+duration_ms,data=train,trControl = train_control,set.seed(30))

summary(model2)

predtest <- predict(model2,newdata =  test,type="response")
table(predtest, test$target) 
mean(predtest == test$target)


```

```{r}
#Random Forest 

train_control <- trainControl(method="cv", number=10)
model3<-randomForest(target~danceability+energy+loudness+acousticness+instrumentalness+valence+tempo+time_signature+duration_ms,data=train,trControl = train_control,set.seed(88))

model3

predtest <- predict(model3,newdata = test,type="response")
table(predtest,test$target) 
mean(predtest == test$target)

```



```{r}
#Clustering to find similar music patterns
cc<-cordata[-16]
head(cc)

#For song recommendation
c<-data.frame(data$danceability,data$energy,data$acousticness,data$instrumentalness,data$valence)
head(c)


```

```{r}

#Elbow Method for finding the optimal number of clusters
set.seed(40)

#For Clustering 
clusterdata1<-scale(cc)

wss <- (nrow(clusterdata1)-1)*sum(apply(clusterdata1,2,var))
  for (i in 2:15) wss[i] <- sum(kmeans(clusterdata1,
                                       centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

#For Song recommendation 

clusterdata11<-scale(c)

wss <- (nrow(clusterdata11)-1)*sum(apply(clusterdata11,2,var))
  for (i in 2:15) wss[i] <- sum(kmeans(clusterdata11,
                                       centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters for recomm",
     ylab="Within groups sum of squares")
```

```{r}

set.seed(40)

#For Clustering 
km1 <- kmeans(cc, 5, nstart = 20)
aggregate(cc, by=list(cluster=km1$cluster), mean)

set.seed(40)
#For Song recommendation 
km2 <- kmeans(c, 4, nstart = 20)
aggregate(c, by=list(cluster=km2$cluster), mean)

```


Cluster 1 has high danceability, energy, medium valence but low acousticness and instrumentalness mostly like mainstream music could have a lot of hits based on previous analysis(Mainstream)
Cluster 2 is similar to cluster 1 but have  more valence and danceablility and  so this cluster could have more mainstream hits 
Cluster 3 has low danceability, energy,valence but very high accousticness and instumentalness (mellow music) 
Cluster 4 is  medium danceability,valence but high energy, instrumentalness and low accousticness (metal like music)
Cluster 5 medium danceability, energy,valence but high acounsticness and low instrumentalness. (Acoustic Music)

```{r}

#

#plot(km,data=c)

#For Clustering 
plotcluster(cc,km1$cluster)

#For Song recommendation 
plotcluster(c,km2$cluster)

```


```{r}
#c$track<-data$track
#c$artist<-data$artist
#c$hit<-data$target
c$clusters<-km2$cluster

km1$size
km2$size

```

```{r}
#Split data in test and train 
data1<-c
d <- createDataPartition(data1$clusters, p=0.75, list=FALSE)
train <- data1[ d,]
test <- data1[-d,]

nrow(train)
nrow(test)
head(test)

```

```{r}
#Multinomial model to predict the clusters
model <- multinom(clusters ~ ., data = train)
summary(model)

```

``` {r}

p<-predict(model, newdata = test, "class")
table(test$clusters, p)
mean(p == test$clusters)


```
Model Accuracy is 99.31


```{r}
#User Input
a<-0.5
data.danceability<-c(a)
data.energy<-c(0.5)
data.acousticness<-c(0.5)
data.instrumentalness<-c(0.5)
data.valence<-c(0.5)
x <- data.frame(data.danceability, data.energy, data.acousticness,data.instrumentalness,data.valence)
x
cluster<-predict(model, newdata = x, "class")
cluster
dataforview<-c
dataforview$track<-data$track
dataforview$artist<-data$artist
dataforview$hit<-data$target
head(dataforview)

b<-sprintf("select track, artist, hit as Mainstream_Hits from dataforview where clusters= %s order by Mainstream_Hits desc limit 10",cluster)

sqldf(b)

```

```{r}
#Split data in train/test  
     cordata$target<-as.factor(cordata$target)
     d <- createDataPartition(cordata$target, p=0.75, list=FALSE)
     train <- cordata[ d,]
      test <- cordata[-d,]

```


```{r}
#UI 

ui <-  fluidPage( 
  h1("SPOTIFY MUSIC DATA"),
  setBackgroundColor(color = c("#F0E68C","#BDB76B"),
                     gradient = "linear"),

  
 tabsetPanel(
   
#TAB 1 


 tabPanel("EDA",
  
    
  sidebarLayout(
   
    sidebarPanel(
       selectInput(inputId = "number",
                      label = " Choose a Number ",
                      choices  =c(5,10,15,20,25,30)),
       
       textOutput("insights"),

      ),
      
      mainPanel(
  
    tabsetPanel( 
      #1
      tabPanel("Popular Artists",           
        h4("Popular Artists of the Decade"),
        plotOutput("popular",width = "700px"),
        tableOutput("avgstat")
      ),
   
     #2
      tabPanel("Songs per Artist",
              
              tableOutput("num")
  
              ),
    

     #3
      tabPanel("Features Correlation",
              
              tableOutput("cor")
  
              )
     
     
        
    )#tabsetPanel
            )#mainPanel
          ),#sidebarLayout
  
      actionButton("display","Average Statistics of Mainstream Music",
          style="background-color: #F0E68C"),
     tableOutput("avg")
  
  
         ),


#TAB 2 

   tabPanel("Hits Prediction", 
          h4("Build Model for Predicting Mainstream Hits"),
          textOutput("modelinfo"),

          
  sidebarLayout(
   
    sidebarPanel(
      
      selectInput(inputId = "variables",
                      label = " Choose Variables",
                      choices  =colnames(train),
                                  selected=c("danceability","energy","loudness","acousticness",
                                  "instrumentalness","valence","tempo","time_signature","duration_ms"),
                                multiple=TRUE),
      
      
      ), #sidebarPanel
    
    mainPanel(
      tabsetPanel(
        #1
        tabPanel("GLM",
                verbatimTextOutput("glm"),
                textOutput("accuracy")
        ),
        #2
        tabPanel("SVM",
                 verbatimTextOutput("svm"),
                 textOutput("accuracy1")
        ),
        #3
        tabPanel("RandomForest",
                 
                 verbatimTextOutput("rf"),
                 textOutput("accuracy2")
                 
        )
      
          )#tabsetPanel
      
    )#mainPanel
  )#sidebarLayout 
          

          ),

  
#TAB 3  

    tabPanel("Genres",
            textOutput("clu"),
          
          sidebarLayout(
   
    sidebarPanel(
       selectInput(inputId = "c",
                      label = " Choose number of Clusters",
                      choices  =c(2,3,4,5,6,7),
                                  selected=4),
       
       selectInput(inputId = "var",
                       label = " Choose variables",
                        choices = colnames(cordata),
                                   selected=c("danceability","energy","acousticness","instrumentalness","valence"),
                                   multiple=TRUE),
       
       textOutput("clus")
                               
    ),#sidebarPanel
    
    mainPanel(
      
      tabsetPanel(
        #1
        tabPanel("Means",
                tableOutput("summary")
              #  textOutput("accuracy")
        ),
        #2
        tabPanel("Plot",
                 plotOutput("clusters")
        ),
        #3
        tabPanel("Model Summary and Accuracy",
                 
                 verbatimTextOutput("modelc"),
                 textOutput("Accuracyc")
                 
        ),
         #4
        tabPanel("Elbow Method",
                 
           plotOutput("elbow")
              )
        
    )#tabsetPanel
      
      
      
    
    )#mainPanel
    )#sidebarLayout
    
    
     ),
          

#TAB 4 
   tabPanel("Song Recommendations", 
          h4(""),
          
          
sidebarLayout(
   
    sidebarPanel(
      
      textOutput("h"),
      sliderInput("danceability","Danceability",min = 0.1, max = 1,value = 0.5,step = 0.1 ),
      sliderInput("energy","Energy",min = 0.1, max = 1,value = 0.5,step = 0.1 ),
      sliderInput("acousticness","Acousticness",min = 0.1, max = 1,value = 0.5,step = 0.1 ),
      sliderInput("instrumentalness","Instrumentalness",min = 0.1, max = 1,value = 0.5,step = 0.1 ),
      sliderInput("valence","Valence",min = 0.1, max = 1,value = 0.5,step = 0.1 )


      

      
                               
    ),#sidebarPanel
    
    mainPanel(
      
      numericInput("vvv"," Enter the number of song suggestions required", value = 10, min = 0, max = 50),
      #selectInput("vvv", "Select a Number",choices  =c(10,15,20,25,30,35,40),
      #                            selected=10),
      textOutput("t"),
      tableOutput("songs")
      
          
    )#mainPanel
  )#sidebarLayout
  
  
  
          ),
 
#TAB 5   
    tabPanel("About", 
          h4("Content"),
          textOutput("data"),
          actionButton("content","Go!",
          onclick ="window.open('https://www.kaggle.com/theoverman/the-spotify-hit-predictor-dataset')", 
          style="background-color: #BDB76B"),
    
         )



  )#tabsetPanel
 
  )#fluidPage
 

```


```{r}
#Server  
server <- function(input, output, session){
  
  #TAB 1
  
  
###
  
  output$num<-renderTable({
    
    
    a<-sprintf("Select artist, Count(track) from data group by artist order by Count(Track) desc limit %s ",input$number)
 popularartist<-sqldf(a)
    
    
  })
  
  

###
  
  output$cor<-renderTable({
    cordata<-data[-1]
     cordata<-cordata[-1]
      cordata<-cordata[-1]
    cor(cordata)
    
  })
  
  
###
  output$insights<-renderText({" The Dataset is balanced. Here you can see a list of popular artist with the maximum number of hit songs in this decade. When you look at the list of songs per artist, you notice that famous artists such as Drake, Taylor swift, etc have all of their songs to be a hit which I believe has to do with their popularity and not just their music.   
There is a clear difference is speechiness in hip hop artists such as Drake compared to pop artists such as taylor swift. Click on the button below to get the average stats of mainstream hit music"})
  
  
###
  observeEvent(input$display,{
    output$avg<- renderTable({
      
      l<-sqldf("select avg(danceability),avg(energy),avg(speechiness),avg(acousticness),avg(instrumentalness),avg(time_signature) from data where Target=1")
      print(l)
    })
   
  })  
  
 
###
output$popular<-renderPlot({
      
a<-sprintf("Select artist, Count(track) from data where Target=1 group by artist order by Count(Track) desc limit %s ",input$number)
 popularartist<-sqldf(a)

 ggplot(data=popularartist,aes(x=reorder(artist,-`Count(track)`),y=`Count(track)`)) +
  geom_bar(stat="identity", fill="#BDB76B")+  geom_text(aes(label=`Count(track)`), vjust=1.6,color="black", size=3.5)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(
   title = 'Artist with maximum number of hit songs from 2010-2019',
    y = 'count of hit songs', x="Artists") 
  

}) 

###
   
  output$avgstat<- renderTable({
    
    b<-sprintf("Select artist, Count(track), avg(speechiness), avg(danceability), avg(valence), avg(energy) from data where Target=1 group by artist order by Count(Track) desc limit %s ",input$number)
    c<-sqldf(b)
      
      print(c)
         })  
    
#TAB 2  
  
  output$modelinfo<-renderText({ "The goal is to predict if a track is going to be a mainstream hit or flop given it's features. Looking at the GLM model we can tell which variables are significant in predicting the target variable. I've build svm and randomforest model using these significant variables as the independent variables and displayed the accuracy of all the models. The highest accuracy achieved is around 85% using k-fold cross validation with Randomforest model. Note: Models may take up to 6 seconds to load. Thank you for your patience."
    })
  
###
  output$glm<-renderPrint({
 
    predictors = paste(input$variables,collapse="+")
    fml = as.formula(sprintf('target ~ %s', predictors ))
    model1 = glm(fml, data=train,family=binomial())
     
    summary(model1)
     
  })

###   
  output$accuracy<-renderText({
    
    
    predictors = paste(input$variables,collapse="+")
    fml = as.formula(sprintf('target ~ %s', predictors ))
    model1 = glm(fml, data=train,family=binomial())
    
    
    predtest <- predict(model1,newdata =  test,type="response")
    t<-table(predtest > 0.5, test$target) 
    e<-sum(diag(t)) / sum(t)
    sprintf("Accuracy of the model is %s",e)

  })
  
###
     output$svm<-renderPrint({
   
    train_control <- trainControl(method="cv", number=10) #k-fold cross validation
       
    predictors = paste(input$variables,collapse="+")
    fml = as.formula(sprintf('target ~ %s', predictors ))
       
    model2<-svm(fml,data=train,trControl = train_control,set.seed(30))
    summary(model2)
    
  })
  
### 
  output$accuracy1<- renderText({
    
    
      train_control <- trainControl(method="cv", number=10) #k-fold cross validation
       
    predictors = paste(input$variables,collapse="+")
    fml = as.formula(sprintf('target ~ %s', predictors ))
       
    model2<-svm(fml,data=train,trControl = train_control,set.seed(30))

      predtest <- predict(model2,newdata =  test,type="response")
     # table(predtest, test$target) 
     x<- mean(predtest == test$target)
      sprintf("Accuracy of the model is %s",x)
      
      
    })
    
###
  
   output$rf<-renderPrint({
   
    train_control <- trainControl(method="cv", number=10) #k-fold cross validation
       
    predictors = paste(input$variables,collapse="+")
    fml = as.formula(sprintf('target ~ %s', predictors ))
       
    model3<-randomForest(fml,data=train,trControl = train_control,set.seed(30))
    model3
    
  })
  
###
   
   output$accuracy2<- renderText({
    
    
      train_control <- trainControl(method="cv", number=10) #k-fold cross validation
       
    predictors = paste(input$variables,collapse="+")
    fml = as.formula(sprintf('target ~ %s', predictors ))
       
    model3<-randomForest(fml,data=train,trControl = train_control,set.seed(20))

      predtest <- predict(model3,newdata =  test,type="response")
     # table(predtest, test$target) 
     x<- mean(predtest == test$target)
      sprintf("Accuracy of the model is %s",x)
      
      
    })
   
   
#TAB 3 
   
###
  output$clu<-renderText({"Here you can find similar music patterns where you can select the optimal number of clusters with the help of the elbow method and select the variables that are important to you. You can view the plot, model summary, accuracy and eblow method that change based on the clusters and variables selected by you. Note: Might take upto 7 seconds to load."})
   
   
   output$clus<-renderText({"The variables I've selected our danceability, energy, acousticness, instrumentalness and valence. The ideal number of clusters are 4.  Cluster 1 has medium danceability, energy,valence but high acounsticness and low instrumentalness. (Acoustic Music). Cluster 2 has low danceability, energy, valence but very high accousticness and instumentalness (mellow music).Cluster 3 has medium danceability,valence but high energy, instrumentalness and low accousticness (metal music).Cluster 4 has high danceability, energy, medium valence but low acousticness and instrumentalness like mainstream music so could have a lot of hits based on previous analysis. Achieved an accuracy of 99.7%"})
   
###   
  output$summary<-renderTable({ 
  
  c<-cordata[,c(input$var)]
   set.seed(40)
  km2 <- kmeans(c, input$c , nstart = 20)
  a<-aggregate(c, by=list(cluster=km2$cluster), mean)
  print(a)

})

###

  output$clusters<- renderPlot({
    
    c<-cordata[,c(input$var)]
    set.seed(40)
    km2 <- kmeans(c, input$c , nstart = 20)
    
    plotcluster(c,km2$cluster)

    })


###

  output$elbow<- renderPlot({
    
     c<-cordata[,c(input$var)]
    set.seed(40)
    km2 <- kmeans(c, input$c , nstart = 20)
    plotcluster(c,km2$cluster)
   
   clusterdata11<-scale(c)

  wss <- (nrow(clusterdata1)-1)*sum(apply(clusterdata11,2,var))
    for (i in 2:15) wss[i] <- sum(kmeans(clusterdata11,
                                       centers=i)$withinss)
  plot(1:15, wss, type="b", xlab="Number of Clusters",
       ylab="Within groups sum of squares")
  
})
  
  
###
   output$modelc<-renderPrint({
     
   c<-cordata[,c(input$var)]
   set.seed(40)
   km2 <- kmeans(c, input$c , nstart = 20)
   c$clusters<-km2$cluster
   
   data1<-c
   d <- createDataPartition(data1$clusters, p=0.75, list=FALSE)
   train <- data1[ d,]
   test <- data1[-d,]


   model <- multinom(clusters ~ ., data = train)
     summary(model)


     
     
   })
  
###
   output$Accuracyc<-renderText({
     
     c<-cordata[,c(input$var)]
   set.seed(40)
   km2 <- kmeans(c, input$c , nstart = 20)
    c$clusters<-km2$cluster
    
    data1<-c
    d <- createDataPartition(data1$clusters, p=0.75, list=FALSE)
    train <- data1[ d,]
    test <- data1[-d,]
     
     model <- multinom(clusters ~ ., data = train)

     p<-predict(model, newdata = test, "class")
     table(test$clusters, p)
     x<-mean(p == test$clusters)
     
     sprintf("Accuracy of the model is %s",x)
     
   })
  
  
  
#TAB 4

###
output$h<-renderText({"Select a value for each based on your mood"})
      
###

output$t<-renderText({"These are a list of tracks that you might like"})
   
   
### 
output$songs<-renderTable({
     
     
    data.danceability<-c(input$danceability)
    data.energy<-c(input$energy)
    data.acousticness<-c(input$acousticness)
    data.instrumentalness<-c(input$instrumentalness)
    data.valence<-c(input$valence)
    
    c<-data.frame(data$danceability,data$energy,data$acousticness,data$instrumentalness,data$valence)
    km <- kmeans(c, 6, nstart = 20)
    c$clusters<-km$cluster
    
    data1<-c
    d <- createDataPartition(data1$clusters, p=0.75, list=FALSE)
    train <- data1[ d,]
    test <- data1[-d,]
    
    model <- multinom(clusters ~ ., data = train)
    
    x <- data.frame(data.danceability, data.energy, data.acousticness,data.instrumentalness,data.valence)
    x
    cluster<-predict(model, newdata = x, "class")
    dataforview<-c
    dataforview$track<-data$track
    dataforview$artist<-data$artist
    dataforview$hit<-data$target
    head(dataforview)
    a<-input$vvv
    b<-sprintf("select track, artist, hit as Mainstream_Hits from dataforview where clusters= %s order by Mainstream_Hits desc limit %s",cluster,a)

    a<-sqldf(b)
    print(a)
    
    
   })
   
   
   
   
   
  
  
#TAB 5 
  
  
  output$data<-renderText({"This is a dataset consisting of features for tracks fetched using Spotify's Web API. This is for the 10s. Starting from Jan 1st, 2010, ending at Dec 31st, 2019.Click on the button below to visit the source and get an understanding of the features."})
  
  
  
  

}
   
```

```{r}

#App
shinyApp(ui,server)

```


```{r}

```




